---
title: "a3_task3_gabe_runte"
author: "Gabe Runte"
date: "11/6/2020"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

```{r, message=FALSE, warning=FALSE}
urchins_raw <- read_csv(here("data","urchins.csv")) %>% 
  clean_names()
```

##Section 1
Size comparisons, purple and red urchins
```{r, message=FALSE, warning=FALSE}
ggplot(urchins_raw, aes(fill= common_name)) +
  geom_histogram(aes(size)) +
  labs(
    x= "Size (cm)",
    y= "Count by size"
  ) +
  facet_wrap(~ common_name) +
  theme(legend.position = "none")

ggplot(urchins_raw, aes(x= common_name, y= size, fill= common_name)) +
  geom_boxplot() +
  labs(
    x= "Species common name",
    y= "Size (cm)"
  ) +
  theme(
    legend.position = "none"
  )
  
```

```{r, message=FALSE, warning=FALSE}
urchins_summary <- urchins_raw %>% 
  group_by(common_name) %>% 
  summarise(mean_size = mean(size), median_size= median(size), sd_size = sd(size), n= n())
  
kable(urchins_summary,
      col.names = c("Common name", "Mean size (cm)", "Median size (cm)", "Standard deviation of sizes (cm)", "Count"))
```

From the histogram, boxplot, and data table, I can see that purple urchins are smaller (average of 4.5cm in diameter)), in general, than red urchins (average of 6.8cm in diameter). Tight boxes in the boxplot, relatively-normally-distributed histograms, and the similarity of mean and median values give me comfort in expecting that this is not a size-skewed subsample in one way or another (the sample size is also huge). Urchins are generally within +/- 50% of the peak of their histogram ditributions, and have reaonable small sd:mean ratios (0.26 and 0.28 for purple and red urchins, respectively), so their spread does not raise any red flags either. 

##Section 2

```{r, message=FALSE, warning=FALSE}

purple_urchins <- urchins_raw %>% 
  filter(common_name=="Purple Urchin") %>% 
  mutate(date= mdy(date)) %>% 
  mutate(year= year(date)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))

purple_summary <- purple_urchins %>% 
  count(site_full, year)
  
ggplot(purple_summary, aes(x=year, y= n, color= site_full, fill= site_full))+
  geom_line()+
  geom_point() +
  labs(
    x="Year",
    y= "Count"
  )

```

